### 2014-09-18 ApTranscriptome_TR.Rmd
working on analysis of Sailfish replicates

@@

```{r load_expression_data, eval=TRUE, echo=FALSE, results='hide'}
# sailfish expression directory
sfexpressionroot <- "results/trinity-full/sailfish-expression-Trinity-cap3-uclust/"

# read in each file using loop
for (j in 1:length(samples)) {
    samp <- samples[j]
    trtval <- as.numeric(str_split_fixed(samp, "-", 2)[2])
    outpre <- gsub("-", "_", samp)
    outname <- paste(outpre, "_quant", sep="")
    read.sailfish.quant(filein=paste(sfexpressionroot, samp, "_quant/quant_bias_corrected.sf", sep=""), outname=outname, samp = samp, trtval = trtval)
}

# combine into long format
A22.TPM <- rbind(A22_0_quant, A22_3_quant, A22_7_quant, A22_10_quant, A22_14_quant, A22_17_quant, A22_21_quant, A22_24_quant, A22_28_quant, A22_31_quant, A22_35_quant, A22_38_quant)
str(A22.TPM)

# convert to data.table
A22.TPM <- data.table(A22.TPM)
setkey(A22.TPM, Transcript, val)

# set "trt" to true values - truncated in file names for convenience
A22.TPM[val==3,val:=3.5]
A22.TPM[val==10,val:=10.5]
A22.TPM[val==17,val:=17.5]
A22.TPM[val==24,val:=24.5]
A22.TPM[val==31,val:=31.5]
A22.TPM[val==38,val:=38.5]
str(A22.TPM)

# Repeat for Ar
Ar.TPM <- rbind(Ar_0_quant, Ar_3_quant, Ar_7_quant, Ar_10_quant, Ar_14_quant, Ar_17_quant, Ar_21_quant, Ar_24_quant, Ar_28_quant, Ar_31_quant, Ar_35_quant, Ar_38_quant)
Ar.TPM <- data.table(Ar.TPM)
setkey(Ar.TPM, Transcript, val)
Ar.TPM[val==3,val:=3.5]
Ar.TPM[val==10,val:=10.5]
Ar.TPM[val==17,val:=17.5]
Ar.TPM[val==24,val:=24.5]
Ar.TPM[val==31,val:=31.5]
Ar.TPM[val==38,val:=38.5]
str(Ar.TPM)
```

Note that expression levels at each temperature treatment are highly correlated between the two colonies.

```{r exp_correlations, echo=FALSE, results='hide'}
temp <- c(0, 3.5, 7, 10.5, 14, 17.5, 21, 24.5, 28, 31.5, 35, 38.5)
cors <- c(round(cor(Ar_0_quant$TPM, A22_0_quant$TPM), 2), round(cor(Ar_3_quant$TPM, A22_3_quant$TPM), 2), round(cor(Ar_7_quant$TPM, A22_7_quant$TPM), 2), round(cor(Ar_10_quant$TPM, A22_10_quant$TPM), 2), round(cor(Ar_14_quant$TPM, A22_14_quant$TPM), 2), round(cor(Ar_17_quant$TPM, A22_17_quant$TPM), 2), round(cor(Ar_21_quant$TPM, A22_21_quant$TPM), 2), round(cor(Ar_24_quant$TPM, A22_24_quant$TPM), 2), round(cor(Ar_28_quant$TPM, A22_28_quant$TPM), 2), round(cor(Ar_31_quant$TPM, A22_31_quant$TPM), 2), round(cor(Ar_35_quant$TPM, A22_35_quant$TPM), 2), round(cor(Ar_38_quant$TPM, A22_38_quant$TPM), 2))

cortable <- cbind(Temperature = temp, Correlation = cors)
```

```{r exp_correlations_table, echo=FALSE, results='asis'}
pandoc.table(cortable, style="rmarkdown", caption = "Correlations between species for gene expression at temperature treatment")
```

@@

```{r combined_TPM, results = 'hide'}
A22.TPM[,colony:="A22"]
Ar.TPM[,colony:="Ar"]
TPM.dt <- rbind(A22.TPM, Ar.TPM)
TPM.dt$colony <- as.factor(TPM.dt$colony)
str(TPM.dt)

setkey(TPM.dt, val)
TPM.dt.sub <- TPM.dt[val != 7] 
unique(TPM.dt.sub$val)
length(unique(TPM.dt.sub$Transcript))
```

@@

# remove from TPM.dt.sub
setkey(TPM.dt.sub, Transcript)
TPM.dt.sub <- TPM.dt.sub[!cont.list]
str(TPM.dt.sub)
length(unique(TPM.dt.sub$Transcript))

>>>>

Preliminary runs of Sailfish revealed that while quantification is highly consistent among runs, small stochastic differences impact which transcripts are found to be thermally-responsive. To address this, I repeated that expression quantification 10 times and retain only the transcripts that are responsive in at least 8 of to 10 replicates.

The gene expression replicates are generated by running the `scripts/sailfish_rep.sh` script, which uses GNU parallel to repeat the `scripts/sailfish_quant.R` script which runs Sailfish for each of the 24 samples.

@@

@@

```{r load_expression_data, eval=TRUE, echo=FALSE, results='hide'}
# sailfish expression directory
sfexpressionroot <- "results/trinity-full/sailfish-expression-Trinity-cap3-uclust/"
# samples
samples <- c("A22-0", "A22-10", "A22-14", "A22-17", "A22-21", "A22-24", "A22-28", "A22-31", "A22-35", "A22-38", "A22-3", "A22-7", "Ar-0", "Ar-10", "Ar-14", "Ar-17", "Ar-21", "Ar-24", "Ar-28", "Ar-31", "Ar-35", "Ar-38", "Ar-3", "Ar-7")

# loop across replicates
reps <- c("rep1", "rep2", "rep3", "rep4", "rep5", "rep6", "rep7", "rep8", "rep9", "rep10")

for(sr in reps) {
  read.sailfish.reps(sr, fd=sfexpressionroot, repout=paste("TPM.dt.", sr, sep=""))
  }
```

For any single replicate, the expression levels at each temperature treatment are highly correlated between the two colonies.

```{r exp_correlations, echo=FALSE, results='hide'}
cortable <- ddply(TPM.dt.rep1, .(val), summarize, cor(TPM[which(colony=="A22")], TPM[which(colony=="Ar")]))
```

```{r exp_correlations_table, echo=FALSE, results='asis'}
colnames(cortable) <- c("Temperature", "r")
pandoc.table(cortable, style="rmarkdown", caption = "Correlations between species for gene expression at temperature treatment", round=2)
```

@@

# setkey of each TPM replicate to 'Transcript'
ls1 <- ls(pattern="TPM")
sapply(mget(ls1), function(x) setkey(x, Transcript))
# remove Transcripts found in contaminant list
sapply(mget(ls1), function(x) x[!cont.list])
